---
title: "HW9_xiuqiqi2"
author: "Xiuqi Qi"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
plot(faithful, pch = 19)
```



# Part (a)

\begin{align*}

Q(\boldsymbol \theta | \boldsymbol \theta^{k}) & = 
E_{\mathbf{Z} | \mathbf{x}, \boldsymbol \theta^{(k)}}[L(\mathbf{x}, \mathbf{z} | \boldsymbol \theta)]\\

& = E_{\mathbf{Z} | \mathbf{x}, \boldsymbol \theta^{(k)}}[\sum_{i=1}^n \mathbf{log} L(\boldsymbol \theta;\mathbf{x_i}, \mathbf{z})]\\

& = \sum_{i=1}^n \sum_{j=1}^2 \mathbf{P}(Z_i =j | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)}) \log L(\boldsymbol \theta_j,x_i) \\


& = \sum_{i=1}^n \sum_{j=1}^2 \mathbf{P}(Z_i =j | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)}) \Big[(1-Z_i)\log\phi_{\boldsymbol \mu_1, \boldsymbol \Sigma_1}(x_i) + Z_i \phi_{\boldsymbol \mu_2, \boldsymbol \Sigma_2}(x_i) + (1-Z_i)\log(1-\tau) + Z_i \log \tau \Big]
\end{align*}





\begin{align*}

\mathbf{P}(Z_i = 1 | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)})

&= \frac{\mathbf{P}(Z_i =1 , \mathbf{x_i} = x_i |\boldsymbol \theta^{(k)})}{\mathbf{P}( \mathbf{x_i} = x_i | \boldsymbol \theta^{(k)})} \\

& = \frac{\mathbf{P}(Z_i =1 , \mathbf{x_i} = x_i |\boldsymbol \theta^{(k)})}{\mathbf{P}(Z_i =1 , \mathbf{x_i} = x_i |\boldsymbol \theta^{(k)}) + \mathbf{P}(Z_i =2 , \mathbf{x_i} = x_i |\boldsymbol \theta^{(k)})} \\

& = \frac{\tau \phi_{\boldsymbol \mu_1^{(k)}, \boldsymbol \Sigma_1^{(k)}}(x_i)}{(1-\tau)\phi_{\boldsymbol \mu_1^{(k)}, \boldsymbol \Sigma_1^{(k)}}(x_i) + \tau \phi_{\boldsymbol \mu_2^{(k)}, \boldsymbol \Sigma_2^{(k)}}(x_i)} \equiv \mathbf{P_i^{(k)}} \\

\mathbf{P}(Z_i = 2 | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)}) 
&= 1- \mathbf{P}(Z_i = 1 | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)}) 
 = 1 - \mathbf{P_i^{(k)}}   
 
\\
\log\phi_{\boldsymbol \mu_j, \boldsymbol \Sigma_j}(x_i) &= -\frac{d}{2} log(2\tau) -\frac{1}{2}log|\boldsymbol \Sigma_j| -\frac{1}{2}(\mathbf{x_i} - \boldsymbol \mu_j)^T \boldsymbol \Sigma_j^{-1} (\mathbf{x_i} - \boldsymbol \mu_j)  \quad j=1,2

\end{align*}

# Part (b)

\begin{align*}
& solve \frac{\partial g(\boldsymbol \theta | \boldsymbol \theta^{(k)})}{\partial \tau}  = 0 \\

\Rightarrow \quad  &\sum_{i=1}^n \sum_{j=1}^2 \mathbf{P}( Z_i =j | \mathbf{x_i} = x_i,\boldsymbol \theta^{(k)}) \Big( \frac{1-Z_i}{1- \tau} + \frac{Z_i}{\tau} \Big) = 0 \\

\Rightarrow \quad &\sum_{i=1}^n \Big( \frac{ \mathbf{p_i^{(k)}} }{\tau} + \frac{1-\mathbf{p_i^{(k)}} }{1 - \tau} \Big) = 0 \\

\Rightarrow \quad &\tau^{(k)} = \frac{\sum_{i=1}^n \mathbf{p_i^{(k)}}}{n} \\

 &\\
&\frac{\partial g(\boldsymbol \theta | \boldsymbol \theta^{(k)})}{\partial \boldsymbol \mu_1}  = 0 \\
\Rightarrow \quad &\sum_{i=1}^n \mathbf{p_i^{(k)}} \boldsymbol \Sigma_1^{-1} (\mathbf{x_i} - \boldsymbol \mu_1) =0 \\
\Rightarrow \quad &\boldsymbol \mu_1^{(k)} =\frac{\sum_{i=1}^n \mathbf{p_i^{(k)}} \mathbf{x_i}}{\sum_{i=1}^n \mathbf{p_i^{(k)}}}\\
\quad
\end{align*}


\begin{align*}
& \frac{\partial g(\boldsymbol \theta | \boldsymbol \theta^{(k)})}{\partial \boldsymbol \Sigma_1^{-1}} = 0 \\

\Rightarrow &\sum_{i=1}^n  \mathbf{p_i^{(k)}} [-\frac{1}{2} \frac{\partial \log|\boldsymbol \Sigma_1|}{\partial \boldsymbol \Sigma_1^{-1}} - \frac{1}{2}\frac{\partial \text{Trace}((\mathbf{x_i} - \boldsymbol \mu_1)^T \boldsymbol \Sigma_1^{-1} (\mathbf{x_i} - \boldsymbol \mu_1))}{ \partial \boldsymbol \Sigma_1^{-1}} ]  =0\\

\Rightarrow  & \sum_{i=1}^n \mathbf{p_i^{(k)}}[\frac{1}{2}\frac{\partial \log|\boldsymbol \Sigma_1^{-1}|}{\partial \boldsymbol \Sigma_1^{-1}}-\frac{1}{2} \frac{\partial \text{Trace}((\mathbf{x_i} - \boldsymbol \mu_1) (\mathbf{x_i} - \boldsymbol \mu_1)^T \boldsymbol \Sigma_1^{-1} )}{ \partial \boldsymbol \Sigma_1^{-1}}] =0 \\

\Rightarrow & \frac{1}{2}\sum_{i=1}^n \mathbf{p_i^{(k)}}[\boldsymbol \Sigma_1 - (\mathbf{x_i} - \boldsymbol \mu_1)^T  (\mathbf{x_i} - \boldsymbol \mu_1)] = 0 \\

\Rightarrow & \boldsymbol \Sigma_1^{(k)} = \frac{\sum_{i=1}^n \mathbf{p_i^{(k)}} (\mathbf{x_i} - \boldsymbol \mu_1^{(k)})  (\mathbf{x_i} - \boldsymbol \mu_1^{(k)})^T}{\sum_{i=1}^n \mathbf{p_i^{(k)}}}
\end{align*}






# Part (c)
## store the final result
```{r}
#detach("package:mixtools", unload=TRUE)
library(mvtnorm)

# this function helps with updating of sigma
sigma_tool = function(index, Ti, x, mu) {
  Ti[index] * (x[index,] - mu) %*% t((x[index,] - mu))
}


# part(c): implementing the algorithm
x = faithful

#initialize everthing
Pi = 0.5
mu1 = c(2, 60)
mu2 = c(4.5, 80)
sigma1 = matrix(c(1, 1, 1 , 160), nrow = 2)
sigma2 = matrix(c(1.5, 15, 15, 180), nrow = 2)
T1 = rep(0.5, nrow(x))
x = as.matrix(x)
colnames(x) = NULL

for (k in 1:10000) {
  # E step
  
  density1 = apply(
    x,
    MARGIN = 1,
    FUN = dmvnorm,
    mean = mu1,
    sigma = sigma1
  )
  density2 = apply(
    x,
    MARGIN = 1,
    FUN = dmvnorm,
    mean = mu2,
    sigma = sigma2
  )
  
  T1_new = Pi * density1 / (Pi * density1 + (1 - Pi) * density2) # T1
  T2_new = 1 - T1_new
  
  # M step
  Pi_new = mean(T1_new)
  
  # update mu1 and mu2
  mu1 = apply(T1_new * x, FUN = sum, MARGIN = 2) / sum(T1_new)
  mu2 = apply(T2_new * x, FUN = sum, MARGIN = 2) / sum(T2_new)
  
  # update sigma1 and sigma2
  sigma_tool = function(index, Ti, x, mu){
    Ti[index] * (x[index,] - mu) %*% t((x[index,] - mu))
  }
  
  sigma1 = lapply(
    1:nrow(x),
    FUN = sigma_tool,
    Ti = T1_new,
    x = x,
    mu = mu1
  )
  sigma1 = Reduce(`+`, sigma1) / sum(T1_new)
  
  sigma2 = lapply(
    1:nrow(x),
    FUN = sigma_tool,
    Ti = T2_new,
    x = x,
    mu = mu2
  )
  sigma2 = Reduce(`+`, sigma2) / sum(T2_new)
  
  if (mean(abs(T1-T1_new)) < 1e-12) {
    print(k)
    print(c(mu1, mu2, Pi_new) )
    break
  }
  else{
    T1 = T1_new
    Pi = Pi_new
    print(c(mu1, mu2, Pi_new) )
  }
  }


T1 = T1_new

result = list(
  "mu1" = mu1,
  "mu2" = mu2,
  "sigma1" = sigma1,
  "sigma2" = sigma2,
  "Pi" = T1
)


```
Algorithm stops at 19th iteration, and gives the result


## store the result of first iteration
```{r}
Pi = 0.5
mu1 = c(2, 60)
mu2 = c(4.5, 80)
sigma1 = matrix(c(1, 1, 1 , 160), nrow = 2)
sigma2 = matrix(c(1.5, 15, 15,180), nrow=2)
T1 = rep(0,nrow(x))
x = as.matrix(x)
colnames(x) = NULL
density1 = apply(x, MARGIN=1, FUN=dmvnorm, mean = mu1, sigma = sigma1)
density2 = apply(x, MARGIN=1, FUN=dmvnorm,mean = mu2, sigma = sigma2)
T1_new = Pi*density1/(Pi*density1+(1-Pi)*density2)
T2_new = 1 - T1_new
Pi_new = mean(T1_new) 
mu1 = apply(T1_new*x, FUN=sum, MARGIN = 2)/sum(T1_new)
mu2 = apply(T2_new*x, FUN=sum, MARGIN = 2)/sum(T2_new)
sigma1 = lapply(1:nrow(x), FUN=sigma_tool, Ti=T1_new, x=x,mu=mu1)
sigma1 = Reduce(`+`, sigma1)/sum(T1_new)
sigma2 = lapply(1:nrow(x), FUN=sigma_tool, Ti=T2_new, x=x,mu=mu2)
sigma2 = Reduce(`+`, sigma2)/sum(T2_new)
result_1st = list("mu1" = mu1,
              "mu2" = mu2,
              "sigma1" = sigma1, 
              "sigma2" = sigma2, 
              "Pi" = T1_new)
```



## plot the result
```{r}
library(mixtools)

mu1 = c(2, 60)
mu2 = c(4.5, 80)
sigma1 = matrix(c(1, 1, 1 , 160), nrow = 2)
sigma2 = matrix(c(1.5, 15, 15, 180), nrow = 2)

addellipse <- function(mu, Sigma, ...)
  {
    ellipse(mu, Sigma, alpha = .05, lwd = 1, ...)
    ellipse(mu, Sigma, alpha = .25, lwd = 2, ...)
  }
  
# plot of initial values
plot(faithful)
addellipse(mu=mu1, Sigma=sigma1, col = "darkorange")
addellipse(mu=mu2, Sigma=sigma2, col = "deepskyblue")

# plot of first iteration
plot(faithful)
addellipse(result_1st$mu1, result_1st$sigma1, col = "darkorange")
addellipse(result_1st$mu2, result_1st$sigma2, col = "deepskyblue")
  
# plot of the final result
plot(faithful)
addellipse(result$mu1, result$sigma1, col = "darkorange")
addellipse(result$mu2, result$sigma2, col = "deepskyblue")
  
  
detach("package:mixtools", unload=TRUE)
```
The above are the plot of initial values, plot of first iteration and the plot of the final result. 
