---
title: "HW6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Question 1

For each method, produce a figure consists of training data, testing data and your fitted curve.
```{r}
library(mlbench)
  data(Ozone)
  
  # Wind will only be used for Q2
  mydata = data.frame("time" = seq(1:nrow(Ozone))/nrow(Ozone), "ozone" = Ozone$V4, "wind" = Ozone$V6)
  
  trainid = sample(1:nrow(Ozone), 250)
  train = mydata[trainid, ]
  test = mydata[-trainid, ]
  par(mfrow=c(1,2))
  
  plot(train$time, train$ozone, pch = 19, cex = 0.5)
  plot(train$wind, train$ozone, pch = 19, cex = 0.5)
  
train = train[order(train$time),]
  
```

## a.
Write your own code (you cannot use bs() or similar functions) to implement a continuous piecewise linear fitting. Pick 3 knots using your own judgment.
```{r}
library(dplyr)
knot = c(1/4, 1/2, 3/4)
train = na.omit(train)

h1 = rep(1, length(train$time))
h2 = train$time
H = cbind(h1,h2)

positive = function(x){
  output = max(0,x)
}


for ( i in 1:length(knot)){
   h = sapply(train$time-knot[i],positive)
   H = cbind(H,h)
}
y = train$ozone
colnames(H)=NULL
H = cbind(y,H)
df = as.data.frame(H)
fit = lm(y~., data = df)

y_hat = predict(fit, newdata =select(df, -y))
lo <- loess(y_hat~train$time)

H1 = select(df, -y)

plot(x=test$time, y=test$ozone, col='green')
points(x=train$time, y=y, col='blue')
lines(x=as.vector(train$time),y=y_hat, col='red')
```

Write your own code to implement a quadratic spline fitting. Your spline should be continuous up to the first derivative. Pick 4 knots using your own judgment.
```{r}
knot = c(0.2,0.4,0.6,0.8)

h1 = rep(1, length(train$time))
h2 = train$time
h3 = h2^2
H = cbind(h1,h2)
H = cbind(H,h3)

positive = function(x){
  output = (max(0,x))^2
}

for ( i in 1:length(knot)){
  h = sapply(train$time-knot[i],positive)
  H = cbind(H,h)
}

y = train$ozone
colnames(H)=NULL
H = cbind(y,H)
df = as.data.frame(H)


fit = lm(y~., data = df)
y_hat = predict(fit, newdata = select(df,-y))
names(y_hat)=NULL


#plot(x=as.vector(train$time),y=y_hat, col='red')
plot(x=test$time, y=test$ozone, col='green')
points(x=as.vector(train$time), y=y, col='blue')
lines(x=as.vector(train$time),y=y_hat, col='red')

```

Produce a same set of basis as (ii) using the bs() function. Note that they do not have to be exactly the same as yours. Verify (figure out how) that the column spaces are the same.
```{r}
library(splines)
library(pracma)
knot = c(0.2,0.4,0.6,0.8)

H = bs(train$time, degree = 2,intercept = TRUE,  knots = knot)

H_old = select(df,-y)
H_old = as.matrix(H_old)

fit = lm( train$ozone~ bs(train$time, degree=2, intercept = TRUE), data=train)
y_hat = predict(fit, newdata = select(df,-y))



plot(x=train$time, y=train$ozone, col='blue')
points(x=test$time, y=test$ozone, col='green')
lines(x=train$time, y=y_hat, col = "red")

sum(rref(H)!=rref(H_old))
```
The row echelon form of both design matrix are the same, so their column spaces are the same. 


Use existing functions to implement a natural cubic spline with 6 knots. Choose your own knots.
```{r}
library(splines)
library(pracma)
knot = seq(0,1,length.out = 8)[2:7]

H = ns(train$time,intercept = TRUE,  knots = knot)

fit = lm(train$ozone~ns(train$time,intercept = TRUE, knots = knot), data=train)
y_hat = predict(fit, newdata = select(df,-y))



plot(x=train$time, y=train$ozone, col='blue')
points(x=test$time, y=test$ozone, col='green')
lines(x=train$time, y=y_hat, col = "red")

```
Use existing functions to implement a smoothing spline. Use the built-in generalized cross-validation method to select the best tuning parameter.

```{r}
train = na.omit(train)
fit = smooth.spline(train$time, train$ozone)
y_hat = predict(fit, train$time)


plot(x=test$time, y=test$ozone, col='green')
points(x=train$time, y=train$ozone, col='blue')
lines(x=train$time, y = y_hat$y, col='red')
```



```{r}
#wind
knot = c(0.2,0.4,0.6,0.8)

h1 = rep(1, length(train$wind))
h2 = train$wind
h3 = h2^2
H = cbind(h1,h2)
H = cbind(H,h3)

positive = function(x){
  output = (max(0,x))^2
}

for ( i in 1:length(knot)){
  h = sapply(train$wind-knot[i],positive)
  H = cbind(H,h)
}

y = train$ozone
colnames(H)=NULL
H = cbind(y,H)
df = as.data.frame(H)

names(y_hat)=NULL

names(H1)=NULL

df = as.data.frame(cbind(H1,H))

fit = lm(y~. ,data=df)




```

Test error
```{r}
test = na.omit(test)
h1 = rep(1, length(test$time))
h2 = test$time
H = cbind(h1,h2)

positive = function(x){
  output = max(0,x)
}


for ( i in 1:length(knot)){
   h = sapply(test$time-knot[i],positive)
   H = cbind(H,h)
}
y = test$ozone
colnames(H)=NULL
H = cbind(y,H)
df = as.data.frame(H)
fit = lm(y~., data = df)

y_hat = predict(fit, newdata =select(df, -y))
lo <- loess(y_hat~test$time)

H = select(df, -y)


knot = c(1/4, 1/2, 3/4)
test = na.omit(test)

h1 = rep(1, length(test$time))
h2 = test$time
H = cbind(h1,h2)

positive = function(x){
  output = max(0,x)
}


for ( i in 1:length(knot)){
   h = sapply(test$time-knot[i],positive)
   H = cbind(H,h)
}
y = test$ozone
colnames(H)=NULL
H = cbind(y,H)
df = as.data.frame(H)
fit = lm(y~., data = df)

y_hat = predict(fit, newdata =select(df, -y))
lo <- loess(y_hat~test$time)

H1 = select(df, -y)


names(y_hat)=NULL
names(H1)=NULL
df_test = as.data.frame(cbind(H1,H))

y_hat = predict(fit, newdata = df_test)

sum((y_hat-test$ozone)^2)/length(y_hat)


```
average prediction error is 28.8845
